```{r ch2libs, include=FALSE}
library(ggplot2)
library(maps)
library(ggthemes)
library(dplyr)
library(dmetric)
library(tidyr)
library(gseddata)
```

# Data {#ch:data}

In this booklet we explain the methodology that we used to obtain one score for Development (D-score) that is globally interprettable. This chapter introduces the data that we use to illustrate our methodology. The data originates from a study by the Global Child Development Group (GCDG) gathering child development data from 16 cohorts.

* Overview of cohorts and instrument (\@ref(sec:gcdgoverview))
* Cohort descriptions (\@ref(sec:cohorts))
* Instruments (\@ref(sec:instruments))

## Overview of cohorts and instruments {#sec:gcdgoverview}

The Global Child Development Group (GCDG) collected data from 16
cohorts. The objective of the study was to develop a population-based
measure to monitor early child development across ages and countries.
The requirements for inclusion were

1. direct assessment of child development;
2. availability of individual milestone scores;
3. spanning ages between 0-5 years; 
4. availability of follow-up measures, at ages 5-10 years.

The effort resulted in a database containing individual data from over 16,000
children from 11 countries. The world map below (Figure: \@ref(fig:worldmap))
colors the countries included in the study. Section \@ref(sec:cohorts) briefly
describes each cohort. Section \@ref(sec:instruments) reviews the measurement
instruments.


```{r worldmap, fig.cap='(ref:worldmap)', fig.keep='all', results='hide', fig.align='center', fig.height=4, fig.width=8}

map.world = map_data(map='world')
map.world$value <- NA
map.world$value <- ifelse(map.world$region =="Bangladesh", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Brazil", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Chile", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="China", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Colombia", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Ecuador", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Ethiopia", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Jamaica", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Madagascar", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="Netherlands", "skyblue1", map.world$value)
map.world$value <- ifelse(map.world$region =="South Africa", "skyblue1", map.world$value)

ggplot() + 
  borders("world", colour = "gray85", fill = "gray80") +
  geom_map(data = map.world, map = map.world, aes(
    map_id = region,
    fill=value
  )) + 
  theme_map() +
  coord_cartesian(ylim = c(-50, 90)) +
  theme(legend.position = "none")

```
(ref:worldmap) Coverage of countries included in the study.

The GCDG data comprises of birth cohorts, impact evaluation studies
and instrument evaluation studies. Table \@ref(tab:cohorts) displays a
brief overview of the instruments used in each sub-study.

Cohort | `by` | `den` | `gri` | `bat` | `vin` | `ddi` |  `bar` | `tep` | `asq` | `sbi`
--------- | --------- | --------- | --------- | --------- | --------- | --------- | --------- | --------- | --------- | ---------
Bangladesh | x |  |  |  |  |  |  |  |  |
Brazil 1 |  | x |  |  |  |  |  |  |  |
Brazil 2 |  |  |  | x |  |  |  |  |  |
Chile 1 | x |  |  |  |  |  |  |  |  |
Chile 2 |  |  |  |  x |  |  |  | x  |  |
China | x |  |  |  |  |  |  |  |  |
Colombia 1 | x |  |  |  |  |  |  |  |  |
Colombia 2 | x | x |  |x  |  |  |  |  | x |
Ecuador |  |  |  |  |  |  | x |  |  |
Ethiopia | x |  |  |  |  |  |  |  |  |
Jamaica 1 |  |  | x |  |  |  |  |  |  |
Jamaica 2 |  |  | x |  |  |  |  |  |  |
Madagascar |  |  |  |  |  |  |  |  |  | x
Netherlands1 |  |  |  |  |  | x |  |  |  |
Netherlands2 |  |  |  |  |  | x |  |  |  |
South Africa | x |  | x |  | x |  |  |  |  |

: (\#tab:cohorts) Overview of instruments administered in the cohorts.



## Cohort descriptions {#sec:cohorts}

<!-- NOTE: Monotonous and repetitive reading. This information is
probably better presented as a table. The more journalistic
alternative would be to highlight the motivation of each study, and
contrast sources and approaches which each other. -->

The cohorts have different designs, age ranges and assessment instruments. In Figure \@ref(fig:design) for each cohort the age range for the development assessments is displayed colored by the developmental instruments that were used. Below the Figure a brief description of each cohort is written. 


```{r design, fig.width=10, fig.keep='all', results='hide', fig.cap = '(ref:design)' }
lean <- gseddata::get_data(cohorts = 40:56)
lean <- clean_data(lean)

design1 <- lean$visit %>% 
  left_join(lean$itm, by = c("subjid", "agedays"))%>%
  mutate(agemos = round(agedays/365.25*12),
         instrument = substr(item, 1, 3)) %>%
  group_by(cohort,agemos,instrument) %>%
  summarise(n = n())

ggplot(design1, aes(x = agemos, y = reorder(cohort, desc(cohort)), group = instrument, color = instrument))+
  geom_point(position = position_dodge(width = 0.5), aes(group = instrument))+
  scale_colour_manual(values = dmetric::get_palette("instrument", "gsed_palettes"), na.value = "grey")+
  xlab("Age (months)") + ylab("")
```
(ref:design) Age range and assessment instrument of included data for each GCDG cohort

The **Bangladesh** study (GCDG-BGD-7MO) was an impact evaluation study
including 1862 children around the age of 18 months. The Bayley Scale
for Infant and Toddler Development-II (`by2`) was administered and
long-term follow-up data were available for the Wechsler Preschool and
Primary Scale of Intelligence (WPPSI) at 5 years [@Tofail2008].

The **Brazil 1** study (GCDG-BRA-1) was a birth-cohort with 3
measurement moments: 644 children at 3 months, 1412 children at 6
months and 1362 children at 12 months. In each round, the Denver
Developmental Screening Test-II (`den`) was administered. Long-term
follow-up data were available for the Wechsler Adult Intelligence
Scale (WAIS) at 18 years [@Victora2006].

The **Brazil 2** study (GCDG-BRA-2) was a birth-cohort with
measurements of 3907 children at 12 months and 3869 children at 24
months. On both occasions the Battelle Development Inventory (`bat`)
was administered [@Moura2010].

The **Chile 1** study (GCDG-CHL-1) was an impact evaluation study of
128 children assessed at 6 months, 1732 children at 12 months and 279
at 18 months. The `by1` was administered at each wave. Long-term
follow-up data were available for the WPPSI at 5-6 years
[@Lozoff2003].

The **Chile 2** study (GCDG-CHL-2) was a birth-cohort of 4869 children
who were administered the Battelle Developmental Inventory (`bat`) at
7-23 months and 9201 children who were administered the Test de
Desarrollo Psicomotor (`tep`) at 24-58 months. For the latter group,
follow-up data were available for the Peabody Picture Vocabulary Test
(PPVT) at 5-6 years [@conteras2015].

The **China** study (GCDG-CHN) was an impact evaluation study that
contained 990 children assessed with the by3 at 18 months
[@Lozoff2016].

The **Colombia 1** study (GCDG-COL-LT45M) was an impact evaluation
study that comprised two waves, wave 1 contained 704 children at 12-24
months and wave 2 631 children at 24-41 months. The `by3` was
administered at each wave. Long-term follow-up data were available for
PPVT at 4-6 years [@Attanasio2014].

The **Colombia 2** study (GCDG-COL-LT42M) was an instrument validation
study where all 1311 children aged 6-42 months were administered the
by3. For a subgroup of 658 children the den, the Ages and Stages
Questionnaire (`asq`) and the `bat` screener were administered as
well. Long-term follow-up data were available for the Fifth Wechsler
Intelligence Scale for Children (WISC-V) and the PPVT
[@Rubio-Codina2016].

In **Ecuador** an impact evaluation study (GCDG-ECU) was performed for
667 children between 0-35 months where the Barrera Moncada (`bar`) was
administered. Long-term follow-up data were available for the PPVT at
5-8 and 9-12 years [@Paxson2010].

The **Ethiopia** study (GCDG-ETH) was a birth-cohort with 193 children
of 12 months in the first wave, 440 children of 30 months at the
second wave and 456 children of 42 months at the third wave. At each
wave the `by3` was administered. Long-term follow-up data were
available for the PPVT at 10-11 years [@Hanlon2009].

The **Jamaica 1** study (GCDG-JAM-LBW) was an impact evaluation study
were the Griffiths Mental Development Scales (`gri`) were administered
to 225 children of 15 months at the first wave, and 218 children of 24
months at the second wave. Long-term follow-up data were available for
WPPSI and PPVT at 6 years [@Walker2004].

The **Jamaica 2** study (GCDG-JAM-STUNTED) was an impact evaluation
study were the `gri` were administered to 159 children at 9-24 months,
21-36 months and at 33-48 months. Long-term follow-up data were
available for `sbi`, Raven’s Coloured Progressive Matrices (Ravens),
and PPVT at 7-8 years and the WAIS at 17-18 years
[@Grantham-McGregor1991].

The **Madagascar** study (GCDG-MDG) was an impact evaluation study
where the `sbi` was administered to 205 children aged 34-42 months.
Long-term follow-up data were available for `sbi` and PPVT at 7-11
years [@Fernald2011].

The **Netherlands 1** study (GCDG-NLD-SMOCC) was an instrument
validation study with a total of 9 waves. At each wave the Dutch
Developmental Instrument (`ddi`) (In the Netherlands known as Van
Wiechenschema) was administered. The first wave included 1985 children
at 1 month, wave 2 1807 children at 2 months, wave 3 1963 children at 3
months, wave 4 1919 children at 6 months, wave 5 1881 children at 9
months, wave 6 1802 children at 12 months, wave 7 1776 children at 15
months, wave 8 1787 children at 18 months, and wave 9 1815 children at
24 months [@herngreen1992].

The **Netherlands 2** study (GCDG-NLD-2) was an instrument validation
study with a total of five waves. At each wave the `ddi` was
administered. Wave 1 included 1016 children at 24 months, wave 2 995
children at 30 months, wave 3 1592 children at 36 months, wave 4 1592
children at 42 months, and wave 5 1024 children at 48 months
[@doove2010].

The **South Africa** study (GCDG-ZAF) was a birth cohort with four
waves. The first wave included 485 children and second wave 275
children, who were assessed at 6 and 12 months, respectively, with the
`by1` and the `gri`. The third wave included 1802 children and the fourth
wave 1614 children, assessed at 24 and 48 months, respectively, with
the Vineland Social Maturity Scale (`vin`) [@Richter2007].




## Instruments {#sec:instruments}

<!-- NOTE: Monotonous and repetitive reading. This information is
probably better presented as a table. Perhaps cluster the instruments
by types (e.g. individual vs screener, one scale vs multiple scales,
binary vs polytomous vs mixed)? Perhaps a simple HOMALS analysis that
displays the most salient features? -->

The **Bayley Scales for Infant and Toddler Development (`by1`,`by2`,
`by3`)** aim to assess infants and toddlers, aged 1-42 months. The
current version is the `by3`, but in GCDG cohorts earlier version were
also used (i.e. `by1` and `by2`)
[@bayley1969][@bayley1993][@bayley2006]. The 326 items of the `by3`
measure three domains: Cognitive items, Motor items (with fine and
gross motor items) and Language items (with expressive and receptive
items). The `by2` contains 277 items and has two additional subscales:
Social-Emotional and Adaptive Behavior. by1 contains 229 items.
<!-- Bayley N. Bayley Scales of Infant Development. New York:
Psychological Corp, 1969. -->
<!-- Bayley, N. The Bayley Scales of Infant Development-II. San
Antonio, TX: Psychological Corporation, 1993. -->
<!-- Bayley N. Bayley Scales of Infant and Toddler Development–Third
Edition: Technical manual. San Antonio, TX: Harcourt Assessment, 2006.
-->

The **Denver Developmental Screening Test (`den`)** is aimed to
identify developmental problems in children up to age six. The 125
dichotomous test items are distributed over the age range from birth
to six years. The Denver covers four domains: personal-social, fine
motor and adaptive, language, and gross motor. The test items are all
directly observed by an examiner and are not dependent on parent
report [@frankenburg1992] [@frankenburg1990].
<!-- Frankenburg WK, Dodds J, Archer P, Shapiro H, Bresnick B. The
Denver II: a major revision and restandardization of the Denver
Developmental Screening Test. Pediatrics 1992; 89: 91–7. -->
<!-- Frankenburg WK, Dodds J, Archer P, et al. The DENVER II Technical
Manual. Denver, CO: Denver Developmental Materials, 1990 -->

The **Griffiths Mental Development Scales (`gri`)** measure the rate
of development in infants and young children in six developmental
areas: loco motor, personal-social, hearing and language, eye and hand
coordination, performance and practical reasoning [@griffiths1967].
<!-- Griffiths R. The Abilities of Babies: A Study in Mental
Measurement. University of London Press, 1967. -->

The **Battelle Developmental Inventory (`bat`)** measures key
developmental skills in children from birth to 7 years, 11 months. The
instrument contains 450 items distributed over five domains: adaptive,
personal-social, communication, motor, and cognitive [@newborg2005].
<!-- Newborg J. Battelle Developmental Inventory-2nd Edition. Rolling
Meadows, IL: Riverside Publishing, 2005. -->

The **Vineland Social Maturity Scale (`vin`)** is a test to assess
social competence. The instrument contains eight subscales that
measure communication skills, general self-help ability, locomotion
skills, occupation skills, self-direction, self-help eating, self-help
dressing and socialization skills [@doll1953].
<!-- Doll EA. The measurement of social competence: A manual for the
Vineland Social Maturity Scale. 1953. -->

The **Dutch Developmental Instrument (`ddi`)** measures early child
development during the ages 0-4 years. The instrument consists of 75
milestone items that are distributed over three domains: fine motor,
adaptive, personal and social behavior; communication; and gross motor
[@schlesinger1981].
<!-- Schlesinger-Was EA. Ontwikkelingsonderzoek van zuigelingen en
kleuters op het consultatiebureau. 1981 -->

The **Barrera Moncada (`bar`)** is a Spanish instrument that measures
the growth and psychological development of children [@barrera1981].
<!-- Barrera Moncada, G. 1981. Crecimiento y desarrollo psicológico
del niño venezolano. Ediciones psicopediátricas. Caracas. -->

The **Test de Desarrollo Psicomotor (`tep`)** is an instrument to
evaluate toddlers aged 2 to 5 years on their development. The items
are distributed over three sub-tests: 16 items assess coordination; 24
items assess language skills and 12 items assess motor skills
[@haeussler1999].
<!-- Haeussler IM, Marchant T. Tepsi: test de desarrollo psicomotor
2-5 años. Eds. Universidad Católica de Chile, 1999 -->

The **Ages and Stages Questionnaire (`asq`)** measures developmental
progress in children between 2 months and 5,5 years. The instrument
assess development in five areas: personal social, gross motor, fine
motor, problem solving and communication. The `asq` contains sets of
30 items in 21 age intervals and is completed by the caregiver
[@squires2009].
<!-- J. Squires, D. Bricker. Ages & Stages Questionnaires®, Third
Edition (ASQ- 3™). A parent-completed child-monitoring system Paul H.
Brookes Publishing Co., Baltimore (2009) -->

The **Stanford Binet Intelligence Scales (`sbi`)** is a cognitive
ability and intelligence test to diagnose developmental deficiencies
in young children. The items are distributed over five subtests: fluid
reasoning, knowledge, quantitative reasoning, visual-spatial processing
and working memory [@roid2003][@hagen1986].
<!-- Roid GH. Stanford–Binet Intelligence Scales, Fifth Edition. WPS
Psychological Tests, 2003. -->
<!-- SB-4 Hagen E, Stattler J, Stanford–Binet Intelligence Scales,
Fourth Edition·Thorndike, 1986. -->


