---
runtime: shiny 
---

# Comparability {#ch:comparability}

This chapter describes challenges and methodology to harmonize development
data from different sources in the following paragraphs:

* Challenge of comparability (\@ref(sec:challenge))
* Bridging instruments by mapping items (\@ref(sec:mapping))
* Overview of promising item mappings (\@ref(sec:viewmapping))



## Challenge of comparability {#sec:challenge}

The ultimate goal of having a global score for development is to compare
children from different populations. In order to do this, the data included in
the GCDG study had to be harmonized. The harmonization protocol described the following:

* The naming scheme for the early developmental test items;
* Every item occupies only one column;
* Items are coded as: 1 – pass; 0 – fail;
* Missings were coded for items not administered to the child because: item not
age appropriate; item below floor for an individual child; or child refused;
* The names of the background variables and harmonized coding for all cohorts.
For example, mothers education was coded in four categories: 0 - none; 1 -
primary; 2 - secondary; 3 - above secondary, height was expressed in centimeters
and weight in kilograms.

To evaluate and analyse the data from the different cohorts as one large
dataset, the cohorts need to be connected. On the one hand this can be
established by using the overlapping instruments in the different cohorts as
connections. In Table \@ref(tab:linkage) the connections between the cohorts via
overlapping instruments are displayed. In the tabpanels the linkages are
displayed by 6 month age blocks. Some cohorts and instruments are not yet
linked. For these cohorts and intruments, bridging was accomplished by mapping
items.


```{r linkage}

knitr::include_app("https://tnochildhealthstatistics.shinyapps.io/GCDG_instrument_linkage/", 
  height = "600px")
```
: (\#tab:linkage) Overview of linkage between the cohorts via overlapping instruments.


## Bridging instruments by mapping items {#sec:mapping}

Items from different instruments can be mapped to each other by similarity. The
by3 was the most frequently administered instrument in the cohorts, so the items
from other instruments were mapped to the by3 items. Mapping was done based on
similar wordings and descriptions in reference manuals by experienced subject
matter experts. Additionally, same-skill items across other instruments were
mapped into groups, if they did not map onto by3 items. Item mappings were
ordered into five domains: fine motor (FM), gross motor (GM), cognitive (COG),
receptive (REC) and expressive (EXP). Figure \@ref(fig:mapping) displays the
connections between the instruments (blue nodes) via the item groups in each
domain.

```{r mapping, echo=FALSE, fig.cap='(ref:mapping)', fig.height=7, fig.keep='all', warning=FALSE, results='hide'}
library(GGally)
library(network)
library(sna)
library(ggplot2)
library(dplyr)

varlist_gcdg <- dmetric::prepare_items(min_cat = 10)
eqset2 <- ddata::get_equates(varlist_gcdg$items)

eqinst <- eqset2 %>% lapply(dscore::rename_gcdg_gsed) 

#instrument vector:
instruments <- unique(gseddata::get_itemtable(dscore::rename_gcdg_gsed(varlist_gcdg$items))$instrument)[-11]


mappings <- lapply(eqinst, function(x){
  y <- instruments %in% substr(x, 1, 3) * 1
  names(y) <- instruments
  y
})

mappings <- data.frame(mappings)

#mapp <- network(mappings, matrix.type = "bipartite")
#ggnet2(mapp, label = TRUE)
#mappings$COG <- rowSums(mappings[, colnames(mappings)[grepl("COG", colnames(mappings))] ])
#mappings$EXP <- rowSums(mappings[, colnames(mappings)[grepl("EXP", colnames(mappings))] ])
#mappings$REC <- rowSums(mappings[, colnames(mappings)[grepl("REC", colnames(mappings))] ])
#mappings$FM <- rowSums(mappings[, colnames(mappings)[grepl("FM", colnames(mappings))] ])
#mappings$GM <- rowSums(mappings[, colnames(mappings)[grepl("GM", colnames(mappings))] ])
#mapp2 <- network(mappings[,c("COG", "EXP", "REC", "GM", "FM")], matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
#ggnet2(mapp2, label = TRUE)
#col = c("actor" = "grey", "event" = "gold")
#ggnet2(mapp2, color = "mode", palette = col, label = TRUE, edge.label = "weights")

col = c("actor" = "lightblue", "event" = "#FC8D62")
mapping_cog <- mappings[, colnames(mappings)[grepl("COG", colnames(mappings))]]
mapping_cog <- mapping_cog[which(rowSums(mapping_cog)>0),]
mapp_cog <-network(mapping_cog, 
                 matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
cog_network<- ggnet2(mapp_cog, 
       label = instruments, 
       color = "mode", 
       alpha = "mode", 
       alpha.palette = c(1, 0.5), 
       palette = col, 
       #shape = "mode",
       #shape.palette = c(15,19), 
       legend.position = "none")+
  ggtitle("Cognitive equates")

col = c("actor" = "lightblue", "event" = "#A6D854")
mapping_exp <- mappings[, colnames(mappings)[grepl("EXP", colnames(mappings))]]
mapping_exp <- mapping_exp[which(rowSums(mapping_exp)>0),]
mapp_exp <-network(mapping_exp,
                 matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
exp_network<- ggnet2(mapp_exp, 
       label = instruments, 
       color = "mode", 
       alpha = "mode", 
       alpha.palette = c(1, 0.5), 
       palette = col, 
       #shape = "mode",
       #shape.palette = c(15,19), 
       legend.position = "none")+
  ggtitle("Expressive equates")

col = c("actor" = "lightblue", "event" = "#8DA0CB")
mapping_fm <- mappings[, colnames(mappings)[grepl("FM", colnames(mappings))]]
mapping_fm <- mapping_fm[which(rowSums(mapping_fm)>0),]
mapp_fm <-network(mapping_fm,
                 matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
fm_network<- ggnet2(mapp_fm, 
       label = instruments, 
       color = "mode", 
       alpha = "mode", 
       alpha.palette = c(1, 0.5), 
       palette = col, 
       #shape = "mode",
       #shape.palette = c(15,19), 
       legend.position = "none")+
  ggtitle("Fine motor equates")

col = c("actor" = "lightblue", "event" = "#E78AC3")
mapping_gm <- mappings[, colnames(mappings)[grepl("GM", colnames(mappings))]]
mapping_gm <- mapping_gm[which(rowSums(mapping_gm)>0),]
mapp_gm <-network(mapping_gm,
                 matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
gm_network<- ggnet2(mapp_gm, 
       label = instruments, 
       color = "mode", 
       alpha = "mode", 
       alpha.palette = c(1, 0.5), 
       palette = col, 
       #shape = "mode",
       #shape.palette = c(15,19), 
       legend.position = "none")+
  ggtitle("Gross motor equates")

col = c("actor" = "lightblue", "event" = "#66C2A5")
mapping_rec <- mappings[, colnames(mappings)[grepl("REC", colnames(mappings))]]
mapping_rec <- mapping_rec[which(rowSums(mapping_rec)>0),]
mapp_rec <-network(mapping_rec,
                  matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
rec_network<- ggnet2(mapp_rec, 
      label = instruments, 
       color = "mode", 
       alpha = "mode", 
       alpha.palette = c(1, 0.5), 
       palette = col, 
       #shape = "mode",
       #shape.palette = c(15,19), 
       legend.position = "none")+
  ggtitle("Receptive equates")

gridExtra::grid.arrange(cog_network, exp_network, gm_network, fm_network, rec_network)

```
(ref:mapping) Connections between the instruments via mapped item groups.

## Overview of promising item mappings {#sec:viewmapping}

```{r plot_p_a_eq, include=FALSE}
library("dmodel")
library("dmetric")
library("gseddata")

theme_set(theme_light())
model <- dmetric::model_lean
lean <- dmetric::gcdg_lean
eq_plots <- plot_p_a_equate(lean, model, passive = TRUE)

#plot_p_a_equate(lean, model, passive = TRUE, file = "C:/Users/eekhouti/Desktop/eq_a_dbook2.pdf")

```

The potential of item mappings is evaluated visually by plotting the probability
to pass items for age for the item groups and compare the curves between
cohorts. When the data shows similar patterns between items and cohorts, the
item mapping can be tested in the modelling phase. In Figure \@ref(fig:poteq),
the probability to pass items for age is displayed for two potential bridges.
The first plot presents a group of four cognitive items that assess the ability
to put a cube or block in a cup or box. The items are administered in different
studies and seem to work similarly. The second plot presents a group of six fine
motor items that assess the ability to build a tower of six cubes or blocks.
These items also seem to work similarly in the different cohorts.

```{r poteq, echo=FALSE, fig.height=7, fig.keep='all', warning=FALSE, results='hide', fig.cap = '(ref:poteq)'}

##promising equates
gridExtra::grid.arrange(
  eq_plots[["COG36"]],
  eq_plots[["FM38"]]
)

```
(ref:poteq) Probability to pass items for age in potential bridges.

When the patterns of the plotted data differ between items and/or cohorts, the
mapping might not be a good one after all. This can be due to differences
between populations or because the items work differently in different
populations (i.e. differential item functioning). The exact cause of the
missmatch can be further investigated in the modelling phase. In Figure
\@ref(fig:badeq), two examples of poor item mappings are presented.

```{r badeq, echo=FALSE,  fig.height=7, fig.keep='all', warning=FALSE, results='hide', fig.cap = '(ref:badeq)'}

##bad equates
gridExtra::grid.arrange(
  eq_plots[["EXP12"]], ## ZA=GRI/BY1; NL=DDI; CL1=BY1
  eq_plots[["COG24"]] ## BR1=den; CO2=den; CL1=BY1; ZA=BY1
)
```
(ref:badeq) Probability to pass items for age in poor bridges.
