---
runtime: shiny 
---

# Comparability {#ch:comparability}

## Challenge of comparability

The ultimate goal of having a global score for development is to compare
children from different populations. In order to do this, the data included in
the GCDG study had to be harmonized. A protocol was written that described a
naming scheme for the early developmental test items and that every item should
only have one column, items should be coded as: 1 – pass; 0 – fail, and that
missings were coded for items not administered to the child because: item not
age appropriate; item below floor for an individual child; or child refused.
Additionally, the names of the background variables were described in the
protocol and to code these variables similarly in all cohorts. For example,
mothers education was coded in four categories: 0 - none; 1 - primary; 2 -
secondary; 3 - above secondary, height was expressed in centimeters and weight
in kilograms.

In order to evaluate and analyse the data from the different cohorts as one
large dataset, the cohorts need to be connected. On the one hand this can be
established by using the overlapping instruments in the different cohorts as
connections. In Table \@ref(tab:linkage) the connections between the cohorts via
overlapping instruments are displayed. In the tabpanels the linkages are
displayed by 6 month age block. As can be observed, some cohorts (and
instruments) are not yet linked. For these cohorts and intruments, bridging was
accomplished by mapping items.


```{r}

knitr::include_app("https://tnochildhealthstatistics.shinyapps.io/GCDG_instrument_linkage/", 
  height = "600px")
```
: (\@tab:linkage) Overview of linkage between the cohorts via overlapping instruments.


## Bridging instruments by mapping items {#sec:mapping}

Items from different instruments can be mapped to each other by similarity. In
the GCDG study, the mapping was done using the Bayley-III items as reference
point. The Bayley-III was the most frequently administered instrument in te
cohorts, so the items from other instruments were mapped to the Bayley-III items
based on similar wordings and descriptions in reference manuals by experienced
subject matter experts. Additionally, same-skill items across other instruments
were mapped into groups, if they did not map onto Bayley-III items. This
resulted in 95 groups of mapped items.


## Overview of promising item mappings {#sec:viewmapping}


```{r plot_p_a_eq, include=FALSE}
library("dmodel")
library("dmetric")
library("gseddata")

theme_set(theme_light())
model <- dmetric::model_lean
lean <- dmetric::gcdg_lean
eq_plots <- plot_p_a_equate(lean, model, passive = TRUE)

#plot_p_a_equate(lean, model, passive = TRUE, file = "C:/Users/eekhouti/Desktop/eq_a_dbook2.pdf")

```

The next step in selecting potential bridges after the item mapping is
performed, is to plot the probability to pass items for age and compare these
between cohorts for mapped items. When the data shows similar patterns between
items and cohorts, the item mapping can be tested in the modelling phase. In
Figure (\@ref(fig:poteq)), the probability to pass items for age is displayed
for two potential bridges. The left plot presents a group of four cognitive
items that assess the ability to put a cube or block in a cup or box. The items
are administered in different studies but seem to work similarly. The plot on
the right presents a group of six fine motor items that assess the ability to
build a tower of six cubes or blocks. These items also seem to work similarly in
the different cohorts they were assessed in.

```{r poteq, echo=FALSE, fig.height=7, fig.keep='all', warning=FALSE, results='hide', fig.cap = '(ref:poteq)'}

##promising equates
gridExtra::grid.arrange(
  eq_plots[["COG36"]],
  eq_plots[["FM38"]]
)

```
(ref:poteq) Probability to pass items for age in potential bridges.

When the patterns of the plotted data differ between items and/or cohorts, the
mapping might not be a good one after all. This can be due to differences
between populations or because the items work differently in different
populations (i.e. differential item functioning). The exact cause of the
missmatch can be further investigated int he modelling phase. In Figure
(\@ref(fig:badeq)), two examples of poor item mappings are presented.

```{r badeq, echo=FALSE,  fig.height=7, fig.keep='all', warning=FALSE, results='hide', fig.cap = '(ref:badeq)'}

##bad equates
gridExtra::grid.arrange(
  eq_plots[["EXP12"]], ## ZA=GRI/BY1; NL=DDI; CL1=BY1
  eq_plots[["COG24"]] ## BR1=den; CO2=den; CL1=BY1; ZA=BY1
)
```
(ref:badeq) Probability to pass items for age in poor bridges.
