```{r ch7libs, include=FALSE}
library(kableExtra)
library(dmetric)
library(gseddata)
library(dplyr)
library(tidyr)
library(ggplot2)
```

# Application I: Tracking a Sustainable Development Goal {#ch:SDGindicator}

> Author: Iris Eekhout

The Sustainable Development Goals (SDG) formulated by the United Nations (UN) set targets to promote prosperity while protecting the planet. One or more indicators quantify the progress towards each target. 

This chapter explores the use of the D-score to monitor the progress of the indicator for healthy child development, SDG 4.2.1. We propose a method to define on-track development and show how the application of this method pans out for the GCDG data. More in detail, the chapter deals with the following topics:

* Estimating SDG 4.2.1 indicator from existing data (\@ref(sec:application1))
* Definition *developmentally on track* (\@ref(sec:references))
* Country-level estimates (\@ref(sec:countrytrack))
* Relation to other estimates (\@ref(sec:otherestimates))

## Estimating SDG 4.2.1 indicator from existing data {#sec:application1}

The [UN Sustainable Development Goals](https://www.un.org/sustainabledevelopment/development-agenda/) form a universal call to action to end poverty, protect the planet and improve the lives and prospects of everyone, everywhere. All UN Member States adopted the 17 Goals in 2015. The SDG 4 target to ensure inclusive and equitable quality education and promote lifelong learning opportunities for all. SDG 4.2 reads as:

> By 2030, ensure that all girls and boys have access to quality early childhood development, care and preprimary education so that they are ready for primary education.

To measure progress, the UN defined indicator 4.2.1 as follows: 

> Proportion of children under 5 years of age who are developmentally on track in health, learning and psychosocial well-being, by sex.

On July 22, 2020, the indicator was [changed](https://unstats.un.org/sdgs/metadata/) into 

> Proportion of children aged 24-59 months who are developmentally on track in health, learning and psychosocial well-being, by sex.

The exclusion of children 0-24 months is at variance with the importance of healthy growth and development during the first 1000 days of life. Indeed, the UN restricted the age range for practical concerns. @loizillon2017 report:

> The initial recommendation was for the ECDI to measure child development from birth–5 years, but the range was restricted to 3–5 years due to time and resource constraints and limited availability of comparable measurement tools for children under age 3.

The careful scientific approach underlying the D-score fills the gap for children aged 0-24 months. Also, the D-score methodology enables extensions to ages beyond 24 months, permits back-calculation of D-scores from existing data, and acts as a linking pin to compare child development from birth onwards.

The cohorts included in the GCDG study represent a wide range of countries and instruments (see Section \@ref(sec:gcdgoverview)). Combining existing data from such a wide range of countries to create the D-score, is undoubtedly challenging, but doable. Although, in all fairness, we note that obtaining accurate comparisons between world-wide populations requires additional representative (existing) data beyond what is available here.

## Definition *developmentally on track* {#sec:references}

For child growth, (global) standards are published by the World Health
Organization (WHO) that are used to monitor growth in children and to define
whether children are *on track*. Growth standards are defined for
weight-for-age, length/height-for age, weight-for-length/height, etc. In 2006,
the methods for constructing a growth reference were reviewed by statisticians
and growth experts that resulted in an advised strategy for constructing growth
standard [@borghi2006]. A similar strategy can be used to construct a reference
standard child development-for-age. This reference standard can then be used to
determine whether children are *developmentally on track*.

In Figure \@ref(fig:dscorerefs), the strategy for constructing standards for
development and to use them to define on track development is visualized in
several steps. By clicking 'Next', the following step will show in the plot. 


```{r dscorerefs, echo=FALSE, out.width="760px", fig.cap='(ref:dscorerefs)' }
#knitr::include_app("https://tnochildhealthstatistics.shinyapps.io/GCDG_references/", 
#  height = "700px")
knitr::include_url("https://iriseekhout.github.io/dbook2tools/gcdgreferences/",
                   height = "800px")
```
(ref:dscorerefs) Illustration of the method to define on-track development


1. In the first step the D-score is plotted by age. 
2. In the second step, this relation is modeled in an LMS model. The goal of an LMS model is to describe the
relation of the D-score by age with age-conditional z-scores. This is done by
three smoothed curves representing the median, coefficient of variation
and the skewness. 
3. In step 3 the centile lines for this model are presented. 
4. The resulting age-standardized scores for development (DAZ), are plotted in step 4.
5. In step 5 the standar deviation lines are drawn to indicate an 1 and 2 standard deviation from the mean.
6. The DAZ is used to define on-track development. So D-scores within 2
standard deviations from the median can be defined as on-track. Scores that are
more than two standard deviation *below* the median, can be defined as off-track
(red dots).

*Note that the current reference is based on the GCDG study. The cohorts that are
currently included are representing a wide range of countries. However, they are
not representative for a global population. The current reference standards are
therefore temporary and should be improved when more data are available.*


## Country-level estimates {#sec:countrytrack}

For each age-D-score combination, a standardized score (DAZ) can be calculated. That way we can
calculate the DAZ in each country that is included in the current GCDG study. By
using the -2SD rule, we can obtain the percentage on-track development for each
country.

```{r, echo = FALSE}
tabtrack <- model_lean$dscore %>%
  mutate(track = ifelse(daz < -2, 0, 1),
         country =  substr(cohort, 6, 8) ) %>%
  group_by(country) %>%
  summarize("on-track"= sum(track, na.rm=TRUE)/sum(!is.na(track)))

kable(tabtrack,
      caption = "Percentage of on-track children per country",
      digits = 2)

```
&nbsp;

## Relation to other estimates {#sec:otherestimates}

In the publication on the global D-score from the GCDG project [@Weber2019], the concurrent, discriminant and predictive validity of the D-score is thoroughly presented and discussed. In this paragraph we mainly focus on stunting, which is commonly used to express impaired growth in children due to nutrition problems. 

### Stunting

The WHO defines stunted as [when the height-for-age is more than two standard deviations below
the WHO Child Growth Standard median](https://www.who.int/nutrition/healthygrowthproj/en/index1.html). In Figure \@ref(fig:stunt) we compare the percentage stunted to the percentage off-track
development in each country. The percentages of stunting differ a lot from the off-track developmental percentages. This indicates that "growth" and "development" are very different constructs. 

```{r stunt, echo=FALSE, fig.cap='(ref:stunt)'}

##change to mean D-score in stunted vs non-stunted children (ref to paper Weber et.al.) But then this plot after that?
#or just load gcdg_lean form dmetric when data in package is updated
sdg_stunt <- 
dmetric::gcdg_lean$visit %>% 
  left_join(dmetric::model_lean$dscore) %>%
  mutate(track = ifelse(.data$daz < -2, 0, 1),
         stunting = ifelse(.data$haz < -2, 1, 0)) %>%
  group_by(ctrycd) %>%
  summarize("off-track" = 1-sum(track, na.rm=TRUE)/sum(!is.na(track)),
            "n_d" = sum(!is.na(track)),
            "stunted" = sum(stunting, na.rm=TRUE)/sum(!is.na(stunting)), 
            "n_stunt" = sum(!is.na(stunting))) %>%
  pivot_longer(cols = c("off-track", "stunted"), names_to = "Measure", values_to = "Percentage")


ggplot(data = sdg_stunt, aes(x = ctrycd, y = Percentage, fill = Measure)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab("Country")


# hazdaz <- dmetric::gcdg_lean$visit %>% 
#   left_join(dmetric::model_lean$dscore) %>%
#   mutate(track = ifelse(.data$daz < -2, 0, 1),
#          stunting = ifelse(.data$haz < -2, 1, 0))%>%
#   group_by(ctrycd) %>%
#   summarize(haz = mean(haz, na.rm = TRUE),
#             daz = mean(daz, na.rm = TRUE),
#             track = mean(track, na.rm = TRUE),
#             stunt = mean(stunting, na.rm = TRUE))
# 
# 
# ggplot(data = hazdaz, aes(x = haz, y = daz, color = ctrycd)) +
#   geom_point() +
#   coord_equal()
# 
# ggplot(data = hazdaz, aes(x = haz, y = daz)) +
#   geom_point() + xlim(-2,2) + ylim(-2,2)+
#   coord_equal()


```
(ref:stunt) Percentage off-track development and stunted per country


However, we can see a relation between stunting and development (D-score). When we compare the Standardized D-score (DAZ) in each country for the stunted versus not-stunted children (Figure \@ref(fig:stuntdaz)). The DAZ for the stunted children is always lower in comparison with the not-stunted children.

```{r stuntdaz, echo=FALSE, fig.width=11, fig.cap= '(ref:stuntdaz)'}
##DAZ in stunted vs non-stunted (as done in Weber paper)
sdg_stuntdaz <- 
dmetric::gcdg_lean$visit %>% 
  left_join(dmetric::model_lean$dscore) %>%
  mutate(stunting = ifelse(.data$haz < -2, "stunted", "not-stunted")) %>%
  group_by(ctrycd, stunting) %>%
  summarize("DAZ" =  mean(daz, na.rm = TRUE)) %>%
  drop_na()

ggplot(sdg_stuntdaz, aes(x = ctrycd, y = DAZ, fill = stunting))+
  geom_bar(stat = "identity", position = "dodge")
```
(ref:stuntdaz) DAZ for stunted versus not-stunted per country


<!--# To validate the off-track development measured by the D-score we can compare the percentage off-track in the studies that administered the `by2` or `by3` using the cut-off for developmental delay (i.e. <70 for `by2`and <78 for `by3` [@https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5940608/]). 
-->


```{r eval=FALSE, include=FALSE}
##cohorts with by3
##cohort with by2
##does not work yet!!!!
#note 14/4/2020: the Bayley scores (and other) are obtianed via study owners and are not part of the data. So we only have the (standardized) scores per cohort, but not specifically related to sdg(off-track). New idea: relate ontrack vs offtrack to the standardized scores per cohort. Hypothesis: when more off-track - lower standardized scores... >>>>> later uitvoeren (staat er nu nog niet in). Eerst even op ch8 richten. 

#join based on subjido & cohort/study. So translate study in ddata gcdg data to cohort in GSEDDATA.
library(dmetric)
library(ddata)
library(dscore)
library(dplyr)
library(tidyr)
library(ggplot2)


gsed_cohort <- unique(dmetric::model_lean$dscore$cohort)
gcdg_aux <- ddata::get_gcdg(items = FALSE, adm=TRUE, aux = TRUE, cov = TRUE)
gcdg_study <- unique(gcdg_aux$study) #brazil2 not in model?

cohort_study <- data.frame(gsed_cohort = gsed_cohort, gcdg_study = gcdg_study[c(4,6,9,10,13,
                                                                                1,5,16,7,8,
                                                                                2,14,15,11,12)])


 gcdg_d_aux <- gcdg_aux %>%
            mutate(subjido = as.character(id), 
                   cohort = cohort_study[match(gcdg_aux$study, cohort_study[,"gcdg_study"]), "gsed_cohort"], 
                   b_total = rowSums(select(.,b_tot_cog, b_tot_lr, b_tot_le, b_tot_mf, b_tot_mg), na.rm =TRUE)) %>% left_join(dmetric::model_lean$dscore, by = c("subjido", "cohort"))
 
 gcdg_d_aux %>% group_by(cohort) %>%
   summarize(cog = mean(b_pc_cog, na.rm=TRUE),
             lang = mean(b_pc_lang, na.rm=TRUE),
             mot = mean(b_pc_mot, na.rm=TRUE),
             fm = mean(b_tot_mf, na.rm=TRUE),
             gm = mean(b_tot_mg, na.rm=TRUE),
             asq_cog = mean(asq_cog, na.rm =TRUE),
             den_mf = mean(den_mf, na.rm = TRUE))
 
 by_track <- gcdg_d_aux %>%
   mutate(track = ifelse(.data$daz < -2, 0, 1), 
          by_d = ifelse(.data$b_total < 70, 1, 0)) %>%
 group_by(country) %>%
  summarize("off-track" = 1-sum(track, na.rm=TRUE)/sum(!is.na(track)),
            "n_d" = sum(!is.na(track)),
            "b_delay" = sum(by_d, na.rm=TRUE)/sum(!is.na(by_d)), 
            "n_b" = sum(!is.na(by_d))) %>%
  pivot_longer(cols = c("off-track", "b_delay"), names_to = "Measure", values_to = "Percentage")


ggplot(data = by_track, aes(x = country, y = Percentage, fill = Measure)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab("Country")

 

 ##vanaf hier verder gaan bij vergeliking met totaal scores op andere measures. 
 
 ##verder deze data gebruiken voor CH8.
 
#select by3 cohorts
by3_COH <- c("GCDG-CHN", "GCDG-COL-LT42M", "GCDG-COL-LT45M", "GCDG-ETH")
#calculate mean of "b_tot_cog"     "b_tot_lr"      "b_tot_le"      "b_tot_mf"      "b_tot_mg"  for bayley total score. 
# add score to dscore data (and visit data for stunting)
#cut score at 70 (or 78) for bayley delay score.

b_vars <- c("b_tot_cog","b_tot_lr","b_tot_le","b_tot_mf","b_tot_mg")

b_scores <- gcdg_lean$aux %>% filter(variable %in% b_vars) %>%
  group_by(subjid)%>% summarize(b_score = mean(value, na.rm=TRUE))



sdg_by <- 
gcdg_lean$visit %>% filter(cohort %in% by3_COH) %>%
  left_join(b_scores) %>%
  left_join(model_lean$dscore) %>%
  mutate(track = ifelse(.data$daz < -2, 0, 1),
                  stunting = ifelse(.data$haz < -2, 1, 0),
         by_d = ifelse(.data$b_score < 70, 1, 0)) %>%
  group_by(ctrycd) %>%
  summarize("off-track" = 1-sum(track, na.rm=TRUE)/sum(!is.na(track)),
            "n_d" = sum(!is.na(track)),
            "stunted" = sum(stunting, na.rm=TRUE)/sum(!is.na(stunting)), 
            "n_stunt" = sum(!is.na(stunting)),
            "b_delay" = sum(by_d, na.rm = TRUE)/sum(!is.na(by_d))) %>%
  pivot_longer(cols = c("off-track", "stunted"), names_to = "Measure", values_to = "Percentage")


ggplot(data = sdg_stunt, aes(x = ctrycd, y = Percentage, fill = Measure)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  xlab("Country")
```